### 2024-04-29 paper/src/con.tex
Update intro
Also moves some stuff to other places

\section{Conclusion}\label{sec:conclusion}
The inability of a particular conceptual framework to match every minute detail of the Galaxy does not invalidate its usefulness. It is, in our view, simplistic to claim any one factor to be the cause of structure in the abundance plane -- especially given the context that much deeper, unobserved structure almost certainly exists. The goal of this work is to argue for a closer examination of a process that has been